---
title: "Penguin Assignment Q4"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have performed a t-test to determine whether the mean bodymass of
gentoo and chinstrap penguins differ significantly from each other and
have produced an overlapping histogram of this data to visually
illustrate this test. My work is presented below in three overall steps:
loading and cleaning the data; performing the t test; and producing and
exporting the plot.

#### Loading and cleaning the data

```{r}
setwd("~/Documents/Year 3/penguin_assessment_Q4") #must be altered to individual's machine

source("functions/libraries.R") #loading in required libraries
source("functions/functions.R") #loading in functions
source("functions/plotting.R") #loading in functions related to plotting

write.csv(penguins_raw, "data_raw/penguins_raw.csv") #preserving a read only version of the raw data

penguin_data_clean <- data_cleaning(penguins_raw) #cleaning data
write.csv(penguin_data_clean, "data_clean/penguin_data_clean.csv") #preserving copy of clean data

removed_empty_bodymass <- remove_empty_bodymass(penguin_data_clean) #remove N/A bodymass values

gentoo_and_chinstrap <- remove_adelie(removed_empty_bodymass) #removing adelie species from dataset
```

#### Performing the T-test

In order to determine what type of T-test should be used one must first
establish whether the data has a normal distribution and whether the two
data sets have equal variance or not.

Shapiro-Wilk test for normality of gentoo data:

H0: data is normal H1: data is not normal Significance level = 0.05

```{r}
#shapiro-wilk test to test normality of gentoo bodymass data
shapiro_wilk_gentoo <- with(gentoo_and_chinstrap, shapiro.test(body_mass_g[species == "Gentoo penguin (Pygoscelis papua)"]))
shapiro_wilk_gentoo
```

p = 0.2336, 0.2336 \> 0.05 therefore accept H0. Data is normal

Shapiro-Wilk test for normality of chinstrap data:

H0: data is normal H1: data is not normal Significance level = 0.05

```{r}
shapiro_wilk_chinstrap <- with(gentoo_and_chinstrap, shapiro.test(body_mass_g[species == "Chinstrap penguin (Pygoscelis antarctica)"])) 
shapiro_wilk_chinstrap 
```

p = 0.5605, 0.5605 \> 0.05 therefore accept H0 Data is normal

Test whether variance is unequal:

H0: variance is equal H1: variance is unequal Significance level = 0.05

```{r}
var.test(body_mass_g ~ species, data = gentoo_and_chinstrap) 
```

p = 0.01559, 0.01559 \< 0.05 therefore reject H0. Variance is not equal

Therefore we cannot use student T test. Welch's t test will be used:

H0: mean bodymass of gentoo and chinstrap penguins are equal H1: mean
bodymass of gentoo and chinstrap penguins are not equal Significance
level = 0.05

```{r}
welch_t_test <- t.test(body_mass_g ~ species, data = gentoo_and_chinstrap, var.equal = FALSE)
welch_t_test
```

p = 2.2e-16, 2.2e-16 \< 0.05 therefore reject H0. Mean bodymass of
gentoo and chinstrap penguins are not equal.

### Producing and exporting the plot

```{r}
plot_bodymass_figure(gentoo_and_chinstrap) #plotting figure

save_bodymass_plot_png(gentoo_and_chinstrap) #exporting figure as pdf

save_bodymass_plot_svg(gentoo_and_chinstrap) #exporting figure as svg
```
